---
output: github_document
---

```{r setup, include=FALSE, message=FALSE, warning=TRUE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  fig.align = 'center', 
  fig.path = "man/figures/")

```

# triplot 

<!-- badges: start -->
[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/triplot)](https://cran.r-project.org/package=triplot)
[![R build status](https://github.com/ModelOriented/triplot/workflows/R-CMD-check/badge.svg)](https://github.com/ModelOriented/triplot/actions?query=workflow%3AR-CMD-check)
[![Codecov test coverage](https://codecov.io/gh/ModelOriented/triplot/branch/master/graph/badge.svg)](https://codecov.io/gh/ModelOriented/triplot?branch=master)
<!-- badges: end -->


## Overview


The `triplot` package provides an instance-level explainer for the groups of explanatory variables called aspect importance. 

Package enables grouping predictors into entities called aspects. Afterwards, it calculates the contribution of those aspects to the prediction for a given observation.

Furthermore, package delivers functionality called `triplot`. It illustrates how the importance of aspects change depending on the size of aspects.

Key functions: 

* `predict_aspects()` for calculating the feature groups importance (called aspects importance) for a selected observation, 
* `predict_triplot()` and `model_triplot()` for summary of automatic aspect importance grouping,
* `group_variables()` for grouping numeric features into aspects.

The `triplot` package is a part of [DrWhy.AI](http://DrWhy.AI) universe. 

## Installation

```{r eval = FALSE}
devtools::install_github("ModelOriented/triplot")
```


## Demo

```{r include=FALSE}
library("triplot")
library("DALEX")

titanic <- titanic_imputed

model_titanic_glm <-
  glm(survived == 1 ~ class + gender + age + sibsp + parch + fare + embarked,
      titanic,
      family = "binomial")

aspects_titanic <-
  list(
    wealth = c("class", "fare"),
    family = c("sibsp", "parch"),
    personal = c("age", "gender"),
    embarked = "embarked"
  )

```

For the `titanic` dataset we built logistic regression model that predicts passanger survival and group features into thematical aspects (`wealth` contains ticket' class and fare, `family` information about family members on board, `personal` includes age and gender and `embarked` remain as aspect with single feature). 
We choose an existing passenger.

```{r }
passenger <- titanic[2,]

passenger

```

We can check that the model prediction for this passenger's survival is very low, it's `r predict(model_titanic_glm, passenger, type = "response")`. 

Let's see which aspects have the biggest influence on it.

We start by building `DALEX` explainer and use it to call `predict_aspects` function. Afterwards, we print and plot function results. We can observe that `wealth` variables have the biggest contribution to the prediction. This contribution is of a negative type. `Family` variables have positive influence on the prediction, but it is many times smaller. Rest of the aspects have very small contribution to the prediction.  

```{r}
explain_titanic <- explain(model_titanic_glm, 
                           data = titanic[, -8],
                           y = titanic$survived == "yes",
                           predict_function = predict,
                           label = "Logistic Regression",
                           verbose = FALSE)

ai_titanic <- predict_aspects(x = explain_titanic, 
                              new_observation = passenger[,-8],
                              variable_groups = aspects_titanic)

print(ai_titanic, show_features = TRUE)

plot(ai_titanic, add_importance = TRUE)
```

## Triplot

`Triplot` is a tool that allows us to go one step further in our understanding of the inner workings of a black box model. 

It illustrates, in one place: 

* the importance of every single feature,
* hierarchical aspects importance, 
* order of grouping features into aspects in `group_variables()`.

We can use it to illustrate the instance level importance of features (using `predict_aspects` function) or to illustrate the model level importance of features (using `model_parts` function from `DALEX` package).

Triplot function can be only used on numerical features. To showcase it, we will choose only such features from `apartments` dataset, build explainer, calculate `model_triplot` object and then plot it.


```{r include=FALSE}

apartments_num <- apartments[,unlist(lapply(apartments, is.numeric))]

new_observation_apartments <- apartments_num[6,-1]

model_apartments <- lm(m2.price ~ ., data = apartments_num)

explain_apartments <- explain(model = model_apartments, 
                              data = apartments_num[, -1],
                              y = apartments_num[, 1],
                              verbose = FALSE)
set.seed(123)
```

We can observe that, at the model level, surface and floor have the biggest contribution. Number of rooms and surface are strongly correlated and together have big influence on the prediction. 

```{r, fig.width=10}
tri_apartments <- model_triplot(explain_apartments)

plot(tri_apartments)
```

Afterwards, we are building triplot for single instance and it's prediction.

We can observe that for the given appartment `surface` has also big, positive influence on prediction. Adding `number of rooms` and then `construction year` to `surface's` aspect, increases its contribution.

We can see notice that `floor` has the smallest influence on the prediction, unlike in model wise analysis.

```{r, fig.width=10}
tri_apartments <- predict_triplot(explain_apartments, 
                                  new_observation = new_observation_apartments)

plot(tri_apartments)
```


## Acknowledgments

Work on this package was financially supported by the 
